Deep Reinforcement Learning for Conservation Decisions

### 2021-05-14 manuscript/manuscript.Rmd
drop bit on conservation policy actions being too simple

Any application of RL can be divided into two components: an *environment* and an *agent*.
An *environment* is typically a computer simulation, though it is possible to use the real world as the RL environment [@robotics].
Over the course of a task, an *agent*, which is often a computer program, continuously interacts with the environment. 
At each time step, the agent observes the current *state* of the environment then performs an *action*; as a result of this action, the environment will transition to a new state and will transmit a numerical *reward* signal to the agent.
The goal of the agent is to learn how to maximize its expected cumulative reward.
The agent learns how to achieve this objective during a period called *training*.
In training, the agent *explores* the available states and actions. 
Once the agent comes across a highly rewarding sequence of states and actions, the agent will reinforce this behavior so that it is more likely for the agent to *exploit* the same high reward trajectory in the future.
Throughout this process, the agent's behavior is codified into what is called a *policy*, which describes what action an agent should take for a given state.

To solve the RL problem, algorithms either take a *model-free* or *model-based* approach. 
The distinction is that *model-free* algorithms do not attempt to learn or use a model of the environment.
Conversely, *Model-based* algorithms either learn or use a model of the environment to achieve the RL objective.
A trade-off between these approaches is that when it is possible to quickly learn a model of the environment or the model is already known, model-based algorithms tend to require much less interaction with the environment to learn good-performing policies [@mbpo].
Yet, frequently, learning a model of the environment is very difficult, and in these cases, model-free algorithms tend to outperform [@CITE].
Since most of the recent focus in RL has gone towards model-free algorithms, we will focus on model-free agents for the rest of the paper.

<!-- Write paragraph on classic approaches then compare to deep RL very broadly-->

@@

An environment is a mathematical function, computer program, or real world experience that takes an agent's proposed *action* as input and returns an observation of its current *state* and an associated *reward* as output.
<!-- ML: what are classical approaches below here? -->
In contrast to classical approaches, there are few restrictions on what comprises a state or action. 
The state may be continuous or discrete, completely or partially observed, single or multidimensional.
For instance, we might represent an ecological community of 10 interacting species with a 10-dimensional continuously valued state space signifying the population density of each species.
The action space can be similarly complex, since an action can be continuous or discrete, single or multidimensional.
For the previous environment, we might set the action space to be continuously valued and 10-dimensional, where each dimension corresponds to the amount of conservation effort aimed at the different species over a time step.   
The main focus of an RL environment, however, is in the environment's transition dynamics and reward function.
The designer of the environment can make the environment follow any transition and reward function provided that both are functions of the current state and action.
This ability allows RL environments to model a broad range of decision making problems.
For example, we can set the transitions to be deterministic or stochastic.
We can also specify the reward function to be *sparse*, whereby a positive reward is only received for state-action pairs that can only be visited after a long sequence of actions, e.g. the end point in a maze.
In other environments, an agent may have to learn to forgo immediate rewards (or even accept an initial negative reward) in order to maximize the net discounted reward over the simulation period, as we illustrate in examples here.
Another important consideration in RL is whether to use an episodic or non-episodic environment.
In episodic environments, there is a finite number of time steps during which the agent interacts with the environment, and for these environments, the designer must specify a stopping condition.
Conversely, with non-episodic environments, the agent can interact with the environment ad infinitum. 
For this paper, we will only consider episodic environments.

@@

@brockman2016 identified (1) the lack of standardization of environments used in publication and (2) the need for better benchmark environments as two core problems limiting more rapid advancement of RL research, and proposed the creation of the OpenAI `gym` framework to address both problems.
The `gym` framework defines a standard interface and methods by which a developer can describe an arbitrary environment in a computer program.
This allows the design and application of software agents which can interact and learn in that environment without knowing anything about the environment's internal details.
The `gym` Python module created by OpenAI also provides over 100 pre-built environments that serve as benchmark problems against which new RL algorithms can be tested.
The `gym` framework is now recognized by all major software frameworks for writing RL algorithms (see Appendix, 'Deep RL Frameworks').
While this library of environments span a range of difficulties from classic control problems that can be solved in closed form to complex three-dimensional robotics simulators, they include no examples of ecological or biological problems.
We use the `gym` framework to turn several existing ecological models into valid environmental simulators that can similarly be used in any RL framework.

## Training and tuning an agent
<!-- Maybe rename this as Training and Tuning an Agent -> how to practically get an agent up and running -->
<!-- Ecologist job to build environment; CS job to build agent 
      Ok so talk about this in very general terms
      selecting an agent, why you might want to use different ones
-->
Training a deep, model-free agent on an environment involves allowing the agent to interact with the environment for thousands or even millions of time steps. 
Over the course of training, the agent will engage in a trial and error process whereby the agent continually updates weights on a collection of neural networks in response to the environment's reward signal, much like how reinforcement learning is believed to work in living creatures [@cite].
Fundamental to this trial and error process is a balancing of explorative and exploitative behavior.
The agent will occasionally need to propose exploratory actions that may result in higher rewards than previously seen. 
Yet, the agent will also need to occasionally exploit the best perceived sequence of actions which may require marginal improvement to reach optimality.
As the exploration-exploitation is currently unsolved, there are a range of approaches that RL algorithms employ.

Model-free algorithms can explore the environment in either an off-policy or on-policy manner. 
The distinction here comes from what policy the agent is using during training.
For off-policy algorithms, the agent is using a policy that can be different than the policy the agent will use during evaluation to select actions; while in on-policy algoirthms, the agent uses the same policy in training as the agent will use during evaluation.
On top of this distinction, agents can learn either a deterministic or stochastic policy.

In practice, the general process for attaining a good performing agent is to tune model hyperparameters then train an agent using the best hyperparameters. 
Hyperparameters refer to the parameters in a RL algorithm that are not directly learned during the training process. <!-- better to define before using in opening sentence.  -->
For example, a weight in the neural network is a model parameter, but the number of hidden layers in the neural network is a hyperparameter.
Since the optimal hyperparameters cannot be found analytically, tuning should be done to find a good-performing set of hyperparameters.
The process of tuning can be automated through standard multi-dimensional optimization approaches: proposing new combinations of hyperparameters, training an agent, and examining the resulting performance.
Since one round of training can be time intensive, hyperparameter tuning, which consists of usually numerous rounds of training, can be particularly time intensive.

@@

@@

There is an ever increasing multitude of reinforcement learning algorithms, and while there are far too many algorithms to discuss in detail in this paper, we will herein classify and touch on some current methods.

@@

```{=html}
<!--

1. Return to theme of RL vs supervised/unsupervised learning.

-->
```
<!-- Something about relevance to both the _theory_ and practice of ecological conservation. (e.g. "strategic" 1-D models have probably had a greater influence, (for better or worse!) on real world policies than complex, detail-specific, precise "tactical" models) -->

## Open Problems

The examples presented here only scrape the surface of possible RL applications to conservation problems.
The examples we have focused on are intentionally quite simple, though it is worth remembering that these very same simple models have a long history of relevance and application in both research and policy contexts.
Despite their simplicity, the optimal strategy is not always obvious before hand, however intuitive it may appear in retrospect.
In the case of the ecosystem tipping point scenario, the optimal strategy is unknown, and the approximate solution found by our RL implementation could almost certainly be improved upon.

Even at it's best, an RL-trained agent can only be as good as the scenarios on which it has been trained.
In these simple examples in which the simulation implements a single model, training is analgous to classical methods which take the model as given [@Marescot2013].
But classical approaches can be difficult to generalize when the underlying model is unknown: alternative models may require a completely different approach.
In contrast, the process of training an RL algorithm on a more complex problem is no different than the simple one: we only need access to a simulation which can generate plausible future states in response to possible actions.
Many of our most realistic ecological models exist only as numerical simulations of lower-level dynamics, such as the individual-based models used in the management of forests and wildfire [@sortie; @firemodel], disease [@disease], marine ecosystems [@ecopath], or global climate change [@gcm].

# Open issues

Deep reinforcement learning has only become an established subfield of AI over the past decade; and, as with most young fields, there are a range of unsolved problems.
In practice, the foremost problem with DRL is instability.
For a range of reasons, including high variance and the exploration vs. exploitation trade-off, model-free and model-based algorithms tend to have unstable performance on environments with non-tabular state-action spaces [cite RL that matters].
For instance, it is not uncommon to observe vastly different behavior when two agents are initialized with different random seeds [cite ibid].
Since model-based and model-free algorithms typically have a large number of hyperparameters, each of which can contribute to training instability, it is practically impossible to identify the source of instability.
The hope of the RL practitioner is that after doing an exhaustive hyperparameter search, the practitioner will find a set of hyperparameters that results in a robust, well-performing agent, but there are no guarantees that hyperparameter tuning will lead to this outcome.
<!-- Kari: what does it mean for two algorithms with similar hyperparameters to perform differently? (As happens in our examples) -->

In the context of conservation, we anticipate that there will be a host of specific challenges to using DRL methods.
Interpretability is likely to be the biggest barrier to adoption.
Although there have been some attempts to increase the transparency of RL agents that use neural network as function approximators [cite], it is still very difficult to explain the behavior of RL-trained agents.
Another challenge is that DRL algorithms are very sample inefficient, and to overcome sample inefficiency for conservation decision-making problems that are often data poor, relying on simulations to train agents will be necessary.
While there are positives to using simulated data, any assumptions or biases that are included in the model will determine agent behavior.

An issue that is central to any applied optimization problem is using the right objective.
There have been many examples in which explicit reward design goes horribly wrong [cite].
And for ecological systems, which are often very complex and impossible to fully observe, it seems that bad reward design is quite likely.
Moreover, specifying a reward function for conservation brings up various ethical concerns -- e.g. who gets to specify the reward function and will the values of the greater community be represented in the reward function.
One recently proposed framework to overcome challenges with explicit reward design is called AI alignment, wherein the agent observes humans, infers the preferences of humans and then acts as to satisfy human preferences.
Yet, in the context of training an agent to say conserve a species, there are questions that could easily arise, foremost will the agent infer that humans don't care about a certain species so acts in a way that leads to that species' extinction?
There is a definite need for exploration of the ethics of using DRL for conservation.

Ultimately, we anticipate that getting DRL to work well on fully detailed conservation decision-making problems will be challenging.
Beyond typically having large state-action spaces for agents to explore, conservation problems tend to have sparse reward structures.
E.g. rarely does an action in conservation lead to immediate rewards; instead, benefits are typically observed after actions have been performed.
Sparse reward structures are challenging for DRL agents, because of the credit assignment problem [cite].
To solve sparse reward environments, many researchers have employed customized algorithms to achieve better performance than standard DRL methods [cite Montezuma].
For conservation, we anticipate that there will likely be a similar pattern, wherein bespoke RL methods will attain better performance than general methods.

-   wild west of AI?
    -can separate RL from having gyms implemented -\> how do you compare different models, -mention that customization is probably the way to go, baselines only go so far

-   AI ethics, transparency

# a vision going forward

-   our public/open source environments and agents, leaderboard
-   the need for collaborative/competitive development both to make better benchmark environments and better agents

>>>>

Any application of RL can be divided into two components: an *environment* and an *agent*.
An *environment* is typically a computer simulation, though it is possible to use the real world as the RL environment [@robotics].
Over the course of an episode, an *agent*, which is often a computer program, continuously interacts with the environment. 
At each time step, the agent performs an *action*, and as a result of this action, the environment will determine the next *state* and numerical *reward* that the agent will observe.
The goal of the agent is to learn how to maximize its expected cumulative reward.
The agent learns how to achieve this objective during a period called *training*.
In training, the agent *explores* the available states and actions. 
Once the agent comes across a highly rewarding sequence of states and actions, the agent will reinforce this behavior so that it is more likely for the agent to *exploit* the same high reward trajectory in future episodes.
Throughout this process, the agent's behavior is codified into what is called a *policy*, which is a mapping of states to actions.

To solve the RL problem, algorithms either take a *model-free* or *model-based* approach. 
The distinction is that *model-free* algorithms do not have access to a model of the environment; yet, *model-based* algorithms use a model of the environment to achieve the RL objective.
A trade-off between these approaches is that when it is possible to quickly learn a model of the environment or the model is already known, model-based algorithms tend to require much less interaction with the environment to learn good-performing policies [@mbpo].
Yet, frequently, learning a model of the environment is very difficult, and in these cases, model-free algorithms tend to outperform [@CITE].
Since most of the recent focus in RL has gone towards model-free algorithms, we will focus on model-free agents for the rest of the paper.

Amongst model-free algorithms, there is divergence around what functions the agent is attempting to learn.
Generally, model-free agents are either learning a policy function, a value function or both.
For context, *value functions* are proxies for how high of a cumulative reward an agent can expect to receive from a given state or state-action pair.
If an agent knows the value function, then the agent can find the optimal action at any state by selecting the action that will maximize the value function in expectation.
The class of algorithms that exclusively try to learn a value function are called *value learning* methods.
*Policy gradient* algorithms exclusively learn the policy.
*Actor-critic* algorithms attempt to learn both a policy function and a value function, whereby the value function, which is learned separately from the policy, is used to inform the agent on the goodness of a selected action.
There is a handful of advantages and disadvantages for each of these approaches that we will not go in depth -- for more, see [@CITE]; yet, the main point of distinction comes from the bias-variance trade-off.
Policy gradients tend to struggle with high variance, value-learning tends to be biased while actor-critic methods balance these two concerns.

@@

An environment is a mathematical function, computer program, or real world experience that takes an agent's proposed *action* as input and returns an observation of it's current *state* and associated *reward* as output.
In contrast to classical approaches, there are few restrictions on what comprises a state or action, at least in principle.
The state may be continuous or discrete, completely or only partially observed, one or many dimensions.
For instance, we might represent a ecological community of 10 interacting species using an environment with 10 continuously valued states representing the possible population densities of each species.
The action space may similarly complex, and the response of the environment to any possible action may be deterministic or stochastic in nature.
The reward will in general depend on both the current state and choice of actions.
An environment must also specify a stopping condition, such as when a particular state is reached or a pre-defined interval of time has elapsed.
In some environments, a reward is only received after the agent successfully completes a potentially long sequence of complex actions, such as solving a maze.
In other environments, an agent may have to learn to forgo immediate reward (or even accept an initial negative reward or cost) in order to maximize the net discounted reward over the simulation period, as we illustrate in examples here.

@@

@brockman2016 identified (1) the lack of standardization of environments used in publication and (2) the need for better benchmark environments as two core problems limiting more rapid advancement of RL research, and proposed the creation of the OpenAI `gym` framework to address both problems.
The `gym` framework defines a standard interface and methods by which a developer can describe an arbitrary environment in a computer program.
This allows the design of software agents which can interact and learn in that environment without knowing anything about its internal details.
The `gym` python module created by OpenAI also provides over 100 pre-built environments that serve as benchmark problems against which new RL algorithms can be tested.
The `gym` framework is now recognized by all major software frameworks for writing RL algorithms (see Appendix, 'Deep RL Frameworks').
While this library of environments span a range of difficulties from classic control problems that can be solved in closed form to complex three-dimensional robotics simulators, they include no examples of ecological or biological problems.
We use the `gym` framework to turn several existing ecological models into valid environmental simulators that can similarly be used in any RL framework.

## Designing and training an agent

The training an agent on an environment involves allowing the agent to experience and respond to repeated simulations from the environment, thousands or even millions of times. 
Over the course of training, weights on the neural network are updated in response to the reward given by the environment, much like reinforcement learning is believed to work in living creatures [@cite].
A neural network can be thought of as a generic function approximator, and these weights as parameters of that function that are estimated from 'data' generated by training.
A typical neural network in RL may have thousands or hundreds of thousands of parameters, giving it the flexibility to learn very complex behaviors.
The overall architecture of the neural network, as well as other parameters such as those governing how an agent learns, how an agent proposes actions, and how the agent discounts future rewards, and so forth are controlled by *hyperparameters*.
To initialize a new, untrained agent, we must specify a handful or so of these hyperparameters, which are held fixed over the course of training.
The particular selection of hyperparameters required depends on which *RL algorithm* our agent will be based.
The RL algorithm specifies precisely how the agent will learn from its environment:
Is the algorithm model-based or model-free? 
Stochastic or deterministic? 
"Off-policy" or "on-policy"? 
Does it use a random noise model to propose actions?
Does it permit a continuous action space or require discrete actions?
The design of RL algorithms has been and remains an area of intense research and rapid development [Table 2].
Open source implementations by professional software development teams often follow swiftly after new algorithms are published [@Ssb3; @tf-agents].
In the following section, we summarize some of the key concepts that make up an RL algorithm, and highlight relevant distinctions between important algorithms today.

<!-- Something about given RL algorithm being able to solve lots of different environments with no pre-specified knowledge. cite examples @AlphaZero and @DQN; but still depends on hyperparameters -->
The process of selecting hyperparameters for a given RL algorithm to maximize it's performance on a given environment is referred to as hyperparameter *tuning*.
The process of tuning can be automated through standard multi-dimensional optimization approaches: proposing new combinations of hyperparameters, training an agent, and examining the resulting performance.
Because training even on a fixed set of hyperparameters can be computationally intensive, tuning must be targeted appropriately.
Considerations in how the environment simulation is implemented, such as the normalization of state and/or action spaces, can significantly improve performance of many RL algorithms on a diverse range of problems without requiring extensive tuning.

@@

Many reinforcement learning algorithms center around estimating value functions, which attempt to quantify the goodness of being in a state or taking an action from a state.
Given that the objective of RL is to maximize the cumulative expected reward, "goodness" refers to the likelihood that a state or state-action pair will result in a high cumulative reward trajectory.
There are 3 value functions that are commonly used across RL algorithms: the state-value function, $V^\pi(s_t)$, the state-action value function, $Q^\pi(s_t, a_t)$, and the advantage function, $A^\pi(s_t, a_t)$.
The state-value function is defined as the expected return from a given state,

$$
  V^\pi(s_t) = \mathbb{E}_{\tau \sim p_\pi(\tau \mid s_t)}\Bigg[\sum_{t' = t}^H \gamma^{t' - t} r(s_t, a_t)\Bigg]
$$

Meanwhile, the state-action-value or quality function, finds the expected return given a state and action,

$$
  Q^\pi(s_t, a_t) = \mathbb{E}_{\tau \sim p_\pi(\tau \mid s_t, a_t)}\Bigg[\sum_{t' = t}^H \gamma^{t' - t} r(s_t, a_t)\Bigg]
$$

And lastly, the advantage function characterizes a relative state-action value in that it quantifies how valuable a state-action pair is over a baseline.
The advantage function can then be succinctly defined as

$$
  A^\pi(s_t, a_t) = Q^\pi(s_t, a_t) - V^\pi(s_t)
$$ 
Learning $Q^\pi$, commonly called the Q-value function, is more powerful than learning the state-value function as Q-values easily prescribe what action to take when in a certain state, $a_t^* = \underset{a_t}{\text{argmax}} \, Q^\pi (s_t, a_t)$.
Conversely, if we know the state-value function, we would have to perform a maximization over an expected value to find the optimal action, $a_t^* = \underset{a_t}{\text{argmax}} \, \mathbb{E}_{s_{t+1} \sim \,T(s_{t+1} \mid s_t, a_t)}[V^\pi(s_{t+1})]$, which is less tractable.
Yet, learning the advantage function is often preferred in practice as advantage estimators favorably have lower variance than state-value and Q-value estimators [cite schulman gae].

@@

There is an ever increasing multitude of reinforcement learning algorithms, and while there are far too many algorithms to discuss in detail in this paper, we will herein classify and touch on some current methods.
RL algorithms can generally be split into 2 groups, model-based and model-free RL.
In model-based RL, the agent is attempting to learn the state transition operator, $T$. <!-- Sometimes the agent is just assuming this is given, right? -->
Once the transition model is known, the agent can then use the transition model to plan out what is the optimal action, often with Model Predictive Control [cite tassa 2012 MPC], or the agent can use the transition model to generate more samples to aid learning [cite Sutton dyna, alphago?]. <!-- For most recent methods focus on an approximate solution, since optimal is computationally impossible.  In contrast most readers will assume this step is trivial, if the transition matrix is known there's nothing more to do but "look" at the predicted transitions and make a choice. -->
In model-free RL, the agent does not attempt to learn a dynamics model and instead tries to either learn a parameterized policy, learn a value function, or try to learn both a parameterized policy and a value function.
The general trade-off between model-free and model-based approaches is that in practice, model-based algorithms are often more sample efficient than model-free algorithms but often fail to surpass model-free algorithms in overall agent performance [cite?]

@@

<!-- Ecology has complex decisions and we lack necessary tools; RL could be it -->
Ecological challenges facing the planet today are complex and outcomes are both uncertain and consequential.
Even our best models and best research will never provide a crystal ball to the future, only better elucidate possible scenarios.
Consequently, that research must also confront the challenge of making robust, resilient decisions in a changing world.
The science of ecological management and quantitative decision-making has a long history [e.g. @Schaefer1954; @Walters1974] and remains an active area of research [@Possingham2006; @Fischer2009; @Polasky2011].
However, the limitations of classical methods such as optimal control frequently constrain applications to relatively simplified models [@Possingham2006], ignoring elements such as spatial or temporal heterogeneity and autocorrelation, stochasticity, imperfect observations, age or state structure or other sources of complexity that are both pervasive and influential on ecological dynamics [@HastingsEncylopedia].
Complexity comes not only from the ecological processes but also the available actions.
Conservation efforts can be implemented in a wide variety of ways, and the most effective policies may vary in both space and time.
While such complex dynamics and large state and action spaces are intractable to classical methods, they are straight forward to implement in simulation environments such as those used to train an RL algorithm.
Neural networks used in deep RL have proven remarkably effective in handling such complexity, particularly when leveraging immense computing resources increasingly available through advances in hardware and software [@gpu_computing].

<!-- Model free is great bc we don't know the true models -->
The rapidly expanding class of Model-free RL algorithms is particularly appealing given the ubiquitous presence of model uncertainty in ecological dynamics.
Rarely do we know the underlying functional forms for ecological processes -- is population recruitment more Beverton-Holt or Ricker?
Methods which must first assume something about the structure or functional form of a process before estimating the corresponding parameter can only ever be as good as those structural assumptions.
Frequently available ecological data is insufficient to distinguish between possible alternative models [@devalpine], or the correct model may be non-identifiable with any amount of data.
Model-free RL approaches offer a powerful way to factor this thorny issue.
The model-free RL agent needs to make no such a priori hypothesis about a particular functional form.
Alternative models can easily be implemented in simulations, allowing an agent to train against both contingencies.
A successful agent will learn a policy that works well in both scenarios.
This is analogous to classical approaches for integrating over parameter or model uncertainty [e.g. @Walters1981; @Ludwig1982],

<!-- Real problems are harder than these examples -->
The examples presented here only scrape the surface of possible RL applications to conservation problems.
The examples we have focused on are intentionally quite simple, though it is worth remembering that these very same simple models have a long history of relevance and application in both research and policy contexts.
Despite their simplicity, the optimal strategy is not always obvious before hand, however intuitive it may appear in retrospect.
In the case of the ecosystem tipping point scenario, the optimal strategy is unknown, and the approximate solution found by our RL implementation could almost certainly be improved upon.
Even at it's best, an RL-trained agent can only be as good as the scenarios on which it has been trained.
In these simple examples in which the simulation implements a single model, training is analgous to classical methods which take the model as given [@Marescot2013].
But classical approaches can be difficult to generalize when the underlying model is unknown: alternative models may require a completely different approach.
In contrast, the process of training an RL algorithm on a more complex problem is no different than the simple one: we only need access to a simulation which can generate plausible future states in response to possible actions.
Many of our most realistic ecological models exist only as numerical simulations of lower-level dynamics, such as the individual-based models used in the management of forests and wildfire [@sortie; @firemodel], disease [@disease], marine ecosystems [@ecopath], or global climate change [@gcm].

<!-- Solving 'real' problems will be hard -->
However, successfully applying RL to complex ecological problems is no easy task.
The flexibility of the approach makes it relatively straightforward to include realistic biology and environmental science in framing an RL problem, but training an RL agent to reach an acceptable level of performance is another task.
As the examples we have presented illustrate, it is important to begin with simple problems, including those for which an optimal strategy is already known.
Such examples provide important benchmarks against which to calibrate the performance, tuning and training requirements of RL,
and can easily be extended into more complex problems once RL agents can master the basics.
Finding acceptable policies even in relatively simple problems may prove beyond the reach of existing algorithms and available computing power.
<!-- something about several technical caveats: instability, interpretability reward design-->

<!-- reason 1: instability-->
One leading problem with deep RL is instability.
For a range of reasons, including high variance [@cite] and the exploration vs. exploitation trade-off [@cite], model-free and model-based algorithms tend to have unstable performance on environments with non-tabular state-action spaces [cite RL that matters].
For instance, it is not uncommon to observe vastly different behavior when two agents are initialized with different random seeds [cite ibid].
Since model-based and model-free algorithms typically have a large number of hyperparameters, each of which can contribute to training instability, it is practically impossible to identify the source of instability.
The hope of the RL practitioner is that after doing an exhaustive hyperparameter search, the practitioner will find a set of hyperparameters that results in a robust, well-performing agent, but there are no guarantees that hyperparameter tuning will lead to this outcome.
<!--CB: I think we can say more about how to diagnose this problem if not fix it.
    Likewise should address the importance of assessing robustness across multiple models/environments, probably at least as important an issue as robustness to nuciance hyperparameters like random seed.
    -->

<!-- reason 2: interpretability --> 
In the context of conservation, we anticipate that there will be a host of specific challenges to using deep RL methods.
Interpretability is likely to be the biggest barrier to adoption. <!-- don't need to rank them -->
Although there have been some attempts to increase the transparency of RL agents that use neural network as function approximators [cite], it is still very difficult to explain the behavior of RL-trained agents.
Another challenge is that DRL algorithms are very sample inefficient, and to overcome sample inefficiency for conservation decision-making problems that are often data poor, relying on simulations to train agents will be necessary.
While there are positives to using simulated data, any assumptions or biases that are included in the model will determine agent behavior.

<!-- reason 3: reward design -->
An issue that is central to any applied optimization problem is using the right objective.
There have been many examples in which explicit reward design goes horribly wrong [cite].
And for ecological systems, which are often very complex and impossible to fully observe, it seems that bad reward design is quite likely.
Moreover, specifying a reward function for conservation brings up various ethical concerns -- e.g. who gets to specify the reward function and will the values of the greater community be represented in the reward function.
One recently proposed framework to overcome challenges with explicit reward design is called AI alignment, wherein the agent observes humans, infers the preferences of humans and then acts as to satisfy human preferences.
Yet, in the context of training an agent to say conserve a species, there are questions that could easily arise, foremost will the agent infer that humans don't care about a certain species so acts in a way that leads to that species' extinction?
There is a definite need for exploration of the ethics of using DRL for conservation.

<!-- reason 4/additional reasons, like sparse rewards -->
Ultimately, we anticipate that getting deep RL to work well on fully detailed conservation decision-making problems will be challenging.
Beyond typically having large state-action spaces for agents to explore, conservation problems tend to have sparse reward structures.
E.g. rarely does an action in conservation lead to immediate rewards; instead, benefits are typically observed after actions have been performed.
Sparse reward structures are challenging for DRL agents, because of the credit assignment problem [cite].
To solve sparse reward environments, many researchers have employed customized algorithms to achieve better performance than standard DRL methods [cite Montezuma].
For conservation, we anticipate that there will likely be a similar pattern, wherein bespoke RL methods will attain better performance than general methods.

<!-- But never fear, hard problems have been solved before!-->
Deep RL is still a very young field, where despite several landmark successes, potential far outstrips practice.
Recent advances in the field have proven the potential of the approach to solve complex problems [@alphaGo2016; @alphaGoZero; @alphazero; @DQN], but typically leveraging large teams with decades of experience in ML and millions of dollars worth of computing power [@alphaGoZero].
Successes have so far been concentrated in applications to games and robotics, not scientific and policy domains, though this is beginning to change [@protein_folding].
Leading developers of new algorithms for deep RL benchmark the performance of their algorithms against such problems in part because simulations of games and robotic movement are readily available through open source frameworks designed to interface with RL algorithms [@gym], and performance can easily be compared against alternatives on public leaderboards [@gym].
Iterative improvements to well-posed public challenges have proven immensely effective in the computer science community in tackling difficult problems, which allow many teams with diverse expertise not only to compete but to learn from each other [@netflix_prize; @imagenet].
Advances in both computational power and algorithmic efficiency continue to reduce costs at a dramatic pace, significantly faster than Moore's law [@openai_costs].
By working to develop similarly well-posed challenges as clear benchmarks, ecology and environmental science researchers may both be better able to replicate that collaborative, iterative success in cracking hard problems.
Such benchmarks may further improve collaboration across disciplines as well, attracting the attention of AI expertise to these key issues facing our planet.
If RL opens the door to greater role for artificial intelligence and engagement of technology firms in the machinery that determines environmental policy, this raises new concerns as well as new opportunities.

<!-- transition from big promise to perils -- Millie's section on risks here -->

<!-- Something about the way forward, point to appendix resources etc -->

<!-- Discussion really doesn't discuss/draw conclusions from the fisheries and tipping points examples! Do we need to or maybe not since it's not about the science-->

<!-- 
The pitfalls of RL-based solutions to global environmental change problems are not only technical, but include issues of ethics and power, especially when the algorithms or data are proprietary.
We discuss not only technical issues that can cause RL solutions to fail in surprising ways, but also potential issues raised by the adoption of more algorithmic approaches to conservation generally.
We emphasize how an open, transparent and reproducible approach can help mitigate some (but not all) of these concerns, and facilitate a more effective interface between teams of researchers from both ecological and computer sciences.
If algorithms are to play a meaningful role in the allocation and exploitation of natural resources from fish to forests, good intentions [@Joppa] may not be sufficient to ensure equitable outcomes.
Ecologists and computer scientists will also need input from experts in governance and ethics to ensure that applications of AI to ecological challenges are fair and just.
